---
title: "Bellabeat Case study"
author: "Sourabh Kumar"
date: "11/15/2021"
output: html_document
---

##Setting up my environment
```{r loading packages, echo=TRUE}
library(tidyverse)
library(readr)
library(lubridate)
library(dplyr)
library(skimr)
```

## Load the csv files
```{r}
dailyActivity_merged <- read_csv("C:/Users/subha/Desktop/Fitabase Data 4.12.16-5.12.16/dailyActivity_merged.csv")
```
```{r}
sleepDay_merged <- read_csv("C:/Users/subha/Desktop/Fitabase Data 4.12.16-5.12.16/sleepDay_merged.csv")
```
##Changed the file name for better readability
```{r}
activity <- dailyActivity_merged
sleep1 <- sleepDay_merged
```
##Data Summary of Both Table
```{r check datasets}
skim_without_charts(activity)
skim_without_charts(sleep)
```
After exploring the data sets, I have noticed that there are some problems with the timestamp data.
I will use the lubridate library's mdy() function to convert date strings as the date elements in daily_activity data are ordered as month, day and year.
```{r}
activity <- activity %>% mutate(ActivityDate = mdy(ActivityDate))
```
```{r}
sleep <-
  sleep %>% 
  mutate(SleepDay = mdy(SleepDay))

```
##Re-check the Data
```{r}
head(activity)
head(sleep)
```
##I will rename both 'ActivityDate' and 'SleepDay' column to 'Date'
```{r}
activity <- 
  activity %>%
  rename(Date = ActivityDate)

sleep <- 
  sleep %>% 
  rename(Date = SleepDay)
```
#from here let`s summarize data
#I will use n_distinct() function to find out the number of respondents in each dataset.
```{r}
n_distinct(activity$Id)
n_distinct(sleep$Id)
```
There are 33 participants in the daily_activity dataset, and only 24 in the sleep-day dataset. This shows that not all users have sleep data.
##Summary Statistics of daily_activity Data:
```{r echo=TRUE}
activity %>% 
  select(TotalSteps,
         TotalDistance,
         SedentaryMinutes, 
         Calories) %>% 
  summary()
```
Based on the summary, I have discovered that the average total steps of the respondents is 7638 per day, the average total distance is 5.490, and the average sedentary time is 991 minutes (16 hours).
The goal of 10,000 steps is the recommended daily step target for healthy adults to achieve health benefits.
##Summary Statistics of sleep_day Data:
```{r echo=TRUE}
sleep %>% 
  select(TotalSleepRecords,
         TotalMinutesAsleep,
         TotalTimeInBed) %>% 
  summary()
```
From this summary, the average time asleep of the respondents is 419.5 minutes (7 hours).
National Sleep Foundation guidelines advise that healthy adults need between 7 and 9 hours of sleep per night.
##Merged both data
```{r}
combined_data <- inner_join(activity, sleep, by = c("Id", "Date"))
n_distinct(combined_data$Id)
```
Now that I have merged the two datasets, there will only be 24 respondents.
